---
title: "RowleyLab2"
author: "Brock Rowley"
date: "1/20/2020"
output: html_document
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(ggridges)
library(RColorBrewer)
```

```{r LoadData}
df <- rio::import(
"https://github.com/fivethirtyeight/data/raw/master/alcohol-consumption/drinks.csv")
```

```{r AlcoholHistogram, fig.width = 12, fig.height = 8}
consume_hist <- ggplot(df, aes(total_litres_of_pure_alcohol)) +
  geom_histogram(
    fill = "#56B4E9",
    color = "white",
    alpha = 0.9,
    bins = 15) + # bins = 5, bins = 25
  labs(x = "Total litres of pure alcohol",
       y = "Count") +
  theme_minimal()
consume_hist
```
At bins = 5, the bin width was too large, smaller features in the distribution
of the data disappear. At bins = 25, the bin width is too small, the histogram
became overly peaky and visually busy and the main trend in the data was
obscured.  

```{r AlcoholDensity, fig.width = 12, fig.height = 8}
consume_dens <- ggplot(df, aes(total_litres_of_pure_alcohol)) +
  geom_density(fill = "#56B4E9",
               bw = 1) + # bw = .5, bw = 1.5
  labs(x = "Total litres of pure alcohol",
       y = "Density") +
  theme_minimal()
consume_dens
```
At bw = .5, the bandwidth is too large, the smaller features in the distribution
of the data disappear. At bw = 1.5, the bandwidth is too small, the density
estimate became overly peaky and visually busy and the main trends in the data
were obscured.  

```{r TypeServings}
df2 <- df %>%
  pivot_longer(
    cols = ends_with("_servings"),
    names_to = "type",
    values_to = "servings"
  )
```

```{r AlcoholRidges, message = FALSE, fig.width = 12, fig.height = 8}
alcohol_dr <- ggplot(df2, aes(servings, type, height = ..density..)) +
  # coord_cartesian(xlim = c(1, 400)) +
  geom_density_ridges_gradient( # this line
    scale = 1,
    fill = "#56B4E9",
    stat = "density",
    trim = TRUE
  ) +
  labs(x = "Total number of servings", # and this one
       y = "Types of alcohol") +
  theme_minimal()

# The two lines noted above are good examples of an inconsistency you're having
# You should pick one or the other, but not use both. Note I slightly formatted
# each, because you had something sort of in-between each

alcohol_dr
```

```{r DensityServingsType, fig.width = 12, fig.height = 8}
alcohol_od <- ggplot(df2, aes(servings)) +
  geom_density(
    aes(fill = type),
    color = "black",
    alpha = 0.5
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal()
alcohol_od
```

```{r WrapServingsType, fig.width = 12, fig.height = 8}
alcohol_wrap <- ggplot(df2, aes(servings)) +
  geom_histogram(aes(fill = type),
                 alpha = 0.9,
                 bins = 10) + # bins = 5, bins = 15
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~type) +
  theme_minimal()

# This is another example where I formatted this one and the one above slightly
# differently. Your styling was sort of in-between again.

alcohol_wrap
```
Double coding the density ridges graph by adding individual colors to the
servings may have provided similar information (visually) as the facet wrap
graphic. I found I needed to work the xlim (trim) for the density ridges graph
because the x-axis extended to -100. My density ridges output seemed to be
communicating negative servings.  
The facet wrap visualization provides an intuitive and highly interpretable
visualization of the drinks data. We clearly see that wine servings outpace
other drink types, in the initial servings count. However, the scale_fill_brewer
is not working as advertised.  
Even though I admit to having limited data interpretation skills, and after
trying several color pallets, density ridges continues to be the most difficult
to interpret for me. This particular density ridges visualization appears (in my
opinion) to have four variables represented.  

```{r ConsumptionType, fig.width = 12, fig.height = 8}
df3 <- df2 %>%
  arrange(
    desc(total_litres_of_pure_alcohol)
  ) %>%
  slice(1:75) %>%
  group_by(country) %>%
  mutate(
    total_servings = sum(servings),
    percent_servings = servings / total_servings
  )

stack <- ggplot(df3, aes(country, percent_servings)) +
  geom_col(
    aes(fill = type),
        alpha = .9
  ) +
  scale_fill_brewer(
    palette = "Paired",
    name = "Alcohol Type",
    breaks = c(
      "wine_servings",
      "spirit_servings",
      "beer_servings"
      ),
    labels = c("Wine", # Note this c() is inconsistent with the breaks c()
               "Spirits",
               "Beer")
    ) +
  coord_flip() +
  scale_y_continuous(
    expand = c(0, 0)
  ) +
  labs(
    y = "Proportion of total alcohol consumed",
    x = NULL,
    title = "Alcohol consumption by type",
    subtitle = "Top 25 countries displayed (according to total liters consumed)"
  ) +
  theme_minimal(
    base_size = 20
  ) +
  theme(
    plot.title.position = "plot",
    legend.position = "bottom"
  )
stack
```